---
title: "Tiered model"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(MASS)
library(Metrics)
```

See `/src/training_tiered.ipynb` for a summary on the tiered modeling strategy.

Load the data for *high* observations:

``` {r, message=FALSE}
X_train_hi <- read_csv("../data/tiered_modeling/X_train_hi.zip")
X_valid_hi <- read_csv("../data/tiered_modeling/X_valid_hi.zip")
y_train_hi <- read_csv("../data/tiered_modeling/y_train_hi.zip")
y_valid_hi <- read_csv("../data/tiered_modeling/y_valid_hi.zip")
```

Load the data for *low* observations:
``` {r, message=FALSE}
X_train_lo <- read_csv("../data/tiered_modeling/X_train_lo.zip")
X_valid_lo <- read_csv("../data/tiered_modeling/X_valid_lo.zip")
y_train_lo <- read_csv("../data/tiered_modeling/y_train_lo.zip")
y_valid_lo <- read_csv("../data/tiered_modeling/y_valid_lo.zip")
```

### Analysis for *high* observations

```{r}
X_train_hi <- X_train_hi %>%
  dplyr::select(-X1)

X_valid_hi <- X_valid_hi %>%
  dplyr::select(-X1)

y_train_hi <- y_train_hi$unacast_session_count

y_valid_hi <- y_valid_hi$unacast_session_count

X_train_hi_model <- X_train_hi %>%
  mutate(target = y_train_hi - 300)
```

Fit a poisson regression model:
``` {r}
poisson_hi <- glm(formula = target ~ ., family = "poisson", data = X_train_hi_model)
```

Print out fitted model:

``` {r}
poisson_hi %>%
  broom::tidy()
```

Residual plot for training data:
```{r}
y_pred_train_hi <- predict(poisson_hi, X_train_hi, type="response") %>%
  broom::tidy() %>%
  mutate(x = x + 300) %>%
  pull(x)

hi_train <- data.frame(y_train_hi, y_pred_train_hi) %>%
  mutate(res = y_train_hi - y_pred_train_hi)

ggplot(hi_train, aes(y_train_hi, res)) + 
  geom_point()
```

Print RMSE for training data:
```{r}
Metrics::rmse(hi_train$y_train_hi, hi_train$y_pred_train_hi)
```

Residual plot for validation data: 
``` {r}
y_pred_valid_hi <- predict(poisson_hi, X_valid_hi, type="response") %>%
  broom::tidy() %>%
  mutate(x = x + 300) %>%
  pull(x)

hi_valid <- data.frame(y_valid_hi, y_pred_valid_hi) %>%
  mutate(res = y_valid_hi - y_pred_valid_hi)

ggplot(hi_valid, aes(y_valid_hi, res)) + 
  geom_point()
```

Print RMSE for validation data:
```{r}
Metrics::rmse(hi_valid$y_valid_hi, hi_valid$y_pred_valid_hi)
```

Identify the anomaly:
``` {r}
outlier <- hi_valid %>%
  arrange(res) %>%
  head(1) %>%
  pull(y_valid_hi)

X_valid_hi_w_target <- X_valid_hi %>%
  mutate(target = y_valid_hi) 

X_valid_hi_w_target %>%
  filter(target == outlier)
```

### Analysis for *low* observations

```{r}
X_train_lo <- X_train_lo %>%
  dplyr::select(-X1)

X_valid_lo <- X_valid_lo %>%
  dplyr::select(-X1)

y_train_lo <- y_train_lo$unacast_session_count

y_valid_lo <- y_valid_lo$unacast_session_count

X_train_lo_model <- X_train_lo %>%
  mutate(target = y_train_lo)
```

Fit a poisson regression model:
``` {r}
poisson_lo <- glm(formula = target ~ ., family = "poisson", data = X_train_lo_model)
```

Print out fitted model:

``` {r}
poisson_lo %>%
  broom::tidy()
```

Residual plot for training data:
```{r}
y_pred_train_lo <- predict(poisson_lo, X_train_lo, type="response") %>%
  broom::tidy() %>%
  pull(x)

lo_train <- data.frame(y_train_lo, y_pred_train_lo) %>%
  mutate(res = y_train_lo - y_pred_train_lo)

ggplot(lo_train, aes(y_train_lo, res)) + 
  geom_point()
```

Print RMSE for training data:
```{r}
Metrics::rmse(lo_train$y_train_lo, lo_train$y_pred_train_lo)
```

Residual plot for validation data: 
``` {r}
y_pred_valid_lo <- predict(poisson_lo, X_valid_lo, type="response") %>%
  broom::tidy() %>%
  pull(x)

lo_valid <- data.frame(y_valid_lo, y_pred_valid_lo) %>%
  mutate(res = y_valid_lo - y_pred_valid_lo)

ggplot(lo_valid, aes(y_valid_lo, res)) + 
  geom_point()
```

Print RMSE for validation data:
```{r}
Metrics::rmse(lo_valid$y_valid_lo, lo_valid$y_pred_valid_lo)
```